Cognitive Principles and Individual Differences \\ in Human Syllogistic Reasoning

### 2019-11-08 paper/paper-old.tex
added paper version as it was submitted

Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
various experimental studies have shown that human reasoning systematically deviates 
from conclusions valid under CL (cf.~\cite{wason:68,byrne:89}).
Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
% as they were mostly theoretical and did not apply to real case studies.
% However, only recently an attempt was made 
% this expertise was not consulted, i.e. there was no communication among the relevant fields.

Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}.

Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions. 
% Only then, the cognitve adequacy of that theory can be compared to others.

Formal methods can offer the required tools to implement such cognitive theories. This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy. In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.

The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans does not only systematically differ from the valid conclusions under classical logic, but from any other proposed cognitive theory (such as the theories proposed in~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).

@@

	\mbox{Mood} & \mbox{natural language} & \mbox{First-order Logic} &

@@

		 \mbox{Figure} & \mbox{Premise 1} & \mbox{Premise 2} \\

@@

		\mbox{1} &  a-b & b-c \\
		\mbox{2} & b-a &  c-b \\
		\mbox{3} & a-b & c-b \\
		\mbox{4} & b-a & b-c \\

@@

concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The valid conclusion \emph{Some artists are not chemists}
under CL, was concluded by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly~\cite{khemlani:2012}.
% \footnote{
% See
%  Given that there are nine different possible conclusions, the chance that a conclusion has been chosen randomly is 1/9 = 11.1 \%. 
%  A binomial test shows that if a conclusion is drawn in
% more than 16\% of the cases by the participants it is unlikely that is has been
% chosen by just random guesses~\cite{khemlani:2012}. }

@@

Usually in these experiments, participants are supposed to give only one answer, thus all three (significant) answers
were given by different participants. 
% The ones who answered \emph{No valid conclusion}
% were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.\footnote{}

@@

As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditionals differently.
This implies that \textit{the} human reasoner does not exist, instead there 
seem to exist various (human) reasoning clusters. The identification of these differences and the specification of these clusters is central 
for the development of an adequate cognitive theory. Only recently approaches have been proposed that account for individual differences (or clusters) in syllogistic reasoning (cf.~\cite{khemlani:2016,declare:2017}).

@@

Here, we focus on a qualitative assessment of the clustering approach in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a more adequate characterization of the individual's reasoning patterns.

\section{Clustering by Principles under WCS} \label{sect:wcsclustering}

According to~\cite{declare:2017}, reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data provided in~\cite{khemlani:2012}. 
The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
Next, the desired models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
 The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.

@@

The suggested computational logic approach for the formal representation and reasoning of these clusters,
is the Weak Completion Semantics (WCS)~\cite{hk:2009a,hk:2009b}
which has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
% WCS is the non-monotonic computational logic approach to consider
% weakly completed logic programs, to compute their least models
% under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.

@@

We introduce an extension of WCS first presented in~\cite{dietz:hoelldobler:pereira:2017}. 
\textit{Contextual (logic) program clauses}
are expressions of the forms $A  \leftarrow  L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ (called
\textit{rules}), $A \leftarrow \top$ (called \textit{facts}), $A \leftarrow \bot$ (called
\textit{negative assumptions})\footnote{Under WCS, the negative
assumption will become $A \leftrightarrow \bot$, which, however, can be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
\textit{unknown assumptions}), where
$A$ is an atom and~$L_i$ with~$1 \leq i \leq m+p$, is a literal.
$A$ is called \textit{head}
and $L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ as well as~$\top, \bot$
and $\udf$,
standing for \textit{true}, \textit{false} and \textit{unknown} respectively, are
called \textit{body} of the corresponding clauses.  A
\textit{contextual program}, denoted by $\mathcal P$,
is a finite set of contextual program clauses.
As we restrict terms to be either constants or variables only, we consider so-called \defname{data logic programs}.
If the clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
% For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
A \defname{ground clause} is a clause that does not contain variables.
$\Ground \CalP$ denotes the set of all its ground clauses.
$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a clause with head $A$.
The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a clause
occurring in } \Ground \CalP\}$.
$A$ is \textit{undefined in $\Ground \CalP$} iff $A$ is not defined in $\Ground \CalP$.
The set of all atoms that are undefined in~$\Ground \CalP$ is denoted by~$\ud(\CalP)$.
Consider the following transformation for a given $\CalP$:
1.\ For each atom~$A$ defined in $\Ground \CalP$, replace
  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
The obtained ground set of equivalences is called the \textit{weak completion}
of~$\CalP$ or~$\WC\CalP$.

% $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
% contains a negative assumption with head $A$, $\Ground \CalP$ does not
% contain an unknown assumption with head $A$, and $\Def(A,\CalP) = \emptyset$.
We omit the word \textit{contextual} when we refer to programs, if not stated otherwise.

@@

We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
extended with the $\cxt$ connective, for which the
corresponding truth values are $\top$, $\bot$ and $\udf$, meaning \textit{true}, \textit{false}
and \textit{unknown}, respectively.
% As it seems,  $\cxt$ corresponds to {\L}ukasiewicz necessity operator~\cite{}, where the difference is, that $\cxt$ is restricted to be applicable to literals only.
A \defname{three-valued interpretation}~$I$ is a mapping from the set of
logical formulas to the set
of truth values~$\{ \true, \bot, \udf\}$, represented as a pair~$I~=~\langle I^\true,I^\bot \rangle$ of two disjoint sets of atoms:  $I^\true =  \{A \mid A \mbox{ is mapped to } \true \mbox{ under } I\}$ and $I^\bot = \{A \mid A \mbox{ is mapped to } \bot \mbox{ under } I \}$.
Atoms which do not occur in~$I^\true \cup I^\bot$ are mapped to~$\udf$.
The truth value of a given formula
under $I$ is determined according to the truth tables in
Table~\ref{tab:3vldluka}.
$I(F) = \top$ denotes that a formula $F$ is mapped to true under $I$.
A \defname{three-valued model} $\CalM$ of~$\CalP$ is a three-valued interpretation such that
$\CalM(A \leftarrow \Body) = \true$ for each $A \leftarrow \Body \in \Ground\CalP$.
Let $I = \langle I^\top, I^\bot \rangle$ and $J = \langle J^\top,
J^\bot \rangle$ be two interpretations.
$I \subseteq J$ iff $I^\top \subseteq J^\top$ and $I^\bot \subseteq J^\bot$. $I$ is the \textit{least model} of $\CalP$
iff for any other model~$J$ of $\CalP$ it holds that $I \subseteq J$.

\begin{table}[t]
\[\small
\begin{array}[c]{@{\hspace{0mm}}c|c}
F & \neg F\\ \midrule
\top & \bot \\
\bot & \top \\
\udf & \udf
\end{array}
\;\
\begin{array}{c|lll}
 \wedge & \top & \udf & \bot \\
\midrule
\top & \top & \udf & \bot \\
\udf & \udf & \udf & \bot \\
\bot & \bot & \bot & \bot
 \end{array}
\;\
\begin{array}{c|lll}
 \vee & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \top & \udf & \udf \\
\bot & \top & \udf & \bot
 \end{array}
\;\
\begin{array}{c|lll}
\leftarrow & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \udf & \top & \top \\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|ccc}
  \leftrightarrow  & \top & \udf & \bot \\
  \midrule
\top & \top & \udf & \bot \\
\udf & \udf & \top & \udf\\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|c}
 L & \cxt(L)\\
\midrule
\top & \top \\
\bot & \bot  \\
\udf & \bot
\end{array}
\]
\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
and \textit{unknown}, respectively.\label{tab:3vldluka}}
\end{table}

@@

\subsection{Reasoning with Respect to the Least Fixed Point of $\Svlp$}\label{sect:lfp}

Consider the following semantic operator~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
\rangle$ be an interpretation. $\Phi_\CalP (I) =
\langle J^\top,J^\bot\rangle$, where
\[
\begin{array}{lcl}
J^\top & = & \{ A \mid A \leftarrow \Body \in \Def(A,\CalP)
\mbox{ and
} \Body \mbox{ is \True under } \langle I^\top,I^\bot \rangle \} \\
J^\bot & = & \{ A \mid \Def(A,\CalP) \neq \emptyset \mbox{ and }\\
&& \rule{7.5mm}{0mm}\Body \mbox{ is \False under } \langle
I^\top,I^\bot \rangle \mbox{ for all }  A \leftarrow \Body \in \Def(A,\CalP)\}.
\end{array}\Shrink
\]
The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
always has a least model under  {\L}ukasiewicz logic,
which can be computed by the least fixed point of~$\Svlp$\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.

The weak completion of contextual
programs might have more than one minimal model, and it is not guaranteed that 
the least fixed point of $\Svlp$ always exists. However,
if these programs do not contain cycles, then there always exists a supported model, which 
corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}.
Let us denote the least fixed point of $\Svlp$ as $\CalM_\CalP$:
As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, we define
$\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
For all programs considered in the sequel, $\CalM_\CalP$ exists.

@@

\subsubsection{Conditional Representation (\conditionals) and License for Inference (\licenses)}\label{sect:condli}
Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
However, it seems that humans understand them defeasibly, they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008},
i.e.\ they might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
explicitly expressed by 
\textit{all $y$ that are not abnormal, are $z$}.
% \\
% Accordingly, $z(X)$ is conjoint with an abnormality predicate: 
% \[\forall X (y(X) \wedge \neg \Ab(X) \rightarrow z(X)).\] 
By default, we assume that \textit{nothing is abnormal}. % with respect to $X$}:
% \[
% \Ab(X) \leftarrow \bot.
% \]

\subsubsection{Existential Import (\import)}

>>>>

Originally, one of the objectives of using logic in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. However, it did not have the expected successes~(cf.~\cite{mcdermott:87}).
Although several alternatives have been proposed, such as non-monotonic or many-valued logics, the emphasis had shifted away from their application to real case studies of human reasoning.
On the other hand, cognitive scientists have extensively investigated human thinking and acquired a great expertise.
However, this expertise was not consulted, i.e. there was no communication among the relevant fields.

Following~\cite{strube:1992}, we argue that a theory that aims at modeling episodes of human reasoning needs to be evaluable according to its cognitive adequacy and should be comparable and compared to other cognitive theories. 
Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}. 
Formal methods can offer the required tools to implement elaborate cognitive theories. This in turn, can allow a new way of communication among the fields, by avoiding ambiguous understanding of the theories, showing their formal correspondences and finally providing an appropriate evaluation method with respect to their cognitive adequacy. In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested principles represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks. 
% Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
% various experimental studies have shown that human reasoning systematically deviates 
% from classical logical conclusions (cf.~\cite{wason:68,byrne:89}).

The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that humans do not only systematically deviate from the assertions in CL, but from any other proposed cognitive theory (such as the theories proposed in~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).

@@

	\mbox{mood} & \mbox{natural language} & \mbox{FOL} &

@@

		& \mbox{premise 1} & \mbox{premise 2} \\

@@

		\mbox{figure 1} &  a-b & b-c \\
		\mbox{figure 2} & b-a &  c-b \\
		\mbox{figure 3} & a-b & c-b \\
		\mbox{figure 4} & b-a & b-c \\

@@

concluded \emph{No chemists are artists} (29\%) and \emph{No valid conclusion} (27\%). The classical logical valid conclusion \emph{Some artists are not chemists}, was concluded 
by only 16\%, which is exactly the threshold, that the answer had not been chosen randomly.\footnote{
 Given that there are nine different possible conclusions, the chance that a conclusion has been chosen randomly is 1/9 = 11.1 \%. 
 A binomial test shows that if a conclusion is drawn in
more than 16\% of the cases by the participants it is unlikely that is has been
chosen by just random guesses~\cite{khemlani:2012}. }

@@

The ones who answered \emph{No valid conclusion}
were not the same ones who answered \emph{No chemists are artists} or \emph{Some artists are not chemists}, and vice versa.

@@

As the above example and recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
This implies that \textit{the} human reasoner does not exist, instead there 
seem to exist various (human) reasoning clusters. The identification of these differences and the specification of these clusters is central 
for the development of an adequate cognitive theory. Only recently approaches have been proposed that account for individual differences (or clusters) in syllogistic reasoning (cf.~\cite{khemlani:2016,declare:2017}).

@@

Here, we focus on a qualitative assessment of the clustering approach in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the cognitive principles. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a new characterization of the individual's reasoning patterns that 
seems to be more adequate.

\section{WCS Clustering} \label{sect:wcsclustering}

\cite{declare:2017} puts forward the hypothesis that reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are (not necessarily valid) assumptions made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data from~\cite{khemlani:2012}. \cite{declare:2017} extended the approach in~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on cognitive principles and heuristic strategies.
After that, 
a translation into logic program clauses was provided, such that for each reasoning clusters, the corresponding logic program
was constructed. Next, the desired models according to the Weak Completion Semantics~\cite{hk:2009a,hk:2009b} were computed with a corresponding semantic operator~\cite{stenning:vanlambalgen:2008}.
After that, the entailed relation between $a$ and $c$ were extracted. If no relation between $a$ and $c$ could be entailed, then \textit{No valid conclusion}
was derived. The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of of multinomial process trees~\cite{Riefer1988}.

@@

As environment for the formalization, we suggest an extension of the
Weak Completion Semantics (WCS). WCS has first been introduced in~\cite{hk:2009a,hk:2009b} 
and has shown to model a wide range of human reasoning tasks~\cite{btg:2015:h}. 
WCS is the non-monotonic computational logic approach to consider
weakly completed logic programs, to compute their least models
under the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}, and to reason with respect to these models or search for explanations by means of (skeptical) abduction~\cite{kakas:etal:93}.

@@

We introduce an extension of WCS first presented in~\cite{dietz:hoelldobler:pereira:2017}. 
\textit{Contextual logic programs} are (data) logic programs extended by the truth-functional operator~$\cxt$, called \textit{context}. \textit{Contextual (logic) program clauses}
are expressions of the forms $A  \leftarrow  L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ (called
\textit{rules}), $A \leftarrow \top$ (called \textit{facts}), $A \leftarrow \bot$ (called
\textit{negative assumptions})\footnote{Under WCS, the negative
assumption will become $A \leftrightarrow \bot$ and, hence, $A$ has to be false.
This can, however, be overwritten by other rules and facts (\textit{defeating} the assumption).} and $A \leftarrow \udf$ (called
\textit{unknown assumptions}), where
$A$ is an atom and the~$L_i$ with~$1 \leq i \leq m+p$ are
literals.
$A$ is called \textit{head}
and $L_1 \land \ldots \land L_m \land
\cxt(L_{m+1}) \land \ldots \land \cxt(L_{m+p})$ as well as~$\top, \bot$
and $\udf$,
standing for \textit{true}, \textit{false} and \textit{unknown} respectively, are
called \textit{body} of the corresponding clauses.  A
\textit{contextual program}, denoted by $\mathcal P$,
is a finite set of contextual program clauses.
As we restrict terms to be either constants or variables only, we consider so-called \defname{data logic programs}.
If the clause contains variables, then they are implicitly universally quantified within the scope of the entire clause. 
For instance, the clause $p(X) \leftarrow q(X)$ represents the formula $(\forall X) (p(X) \leftarrow q(X))$. Therefore, a clause never contains free variables.
A \defname{ground clause} is a clause that does not contain variables.
$\Ground \CalP$ denotes the set of all its ground clauses.
$A$ is \textit{defined in $\Ground \CalP$}  iff $\Ground \CalP$ contains a rule, a fact 
or an assumption with head $A$.
The \textit{definition of $A$ in $\Ground \CalP$} is defined as $\Def(A,\CalP)
= \{ A \leftarrow \Body \mid A \leftarrow \Body \mbox{ is a rule, a fact or an assumption
occurring in } \Ground \CalP\}$.
$A$ is \textit{undefined in $\Ground \CalP$} iff $A$ is not defined in $\Ground \CalP$.
The set of all atoms that are undefined in~$\Ground \CalP$ is denoted by~$\ud(\CalP)$.
% $\neg A$ is \textit{assumed in $\Ground \CalP$} iff $\Ground \CalP$
% contains a negative assumption with head $A$, $\Ground \CalP$ does not
% contain an unknown assumption with head $A$, and $\Def(A,\CalP) = \emptyset$.
We omit the word \textit{contextual} when we refer to programs, if not stated otherwise.

@@

We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
extended with the $\cxt$ connective, for which the
corresponding truth values are $\top$, $\bot$ and $\udf$, meaning \textit{true}, \textit{false}
and \textit{unknown}, respectively.
A \defname{three-valued interpretation}~$I$ is a mapping from the set of
logical formulas to the set
of truth values~$\{ \true, \bot, \udf\}$, represented as a pair~$I~=~\langle I^\true,I^\bot \rangle$ of two disjoint sets of atoms:  $I^\true =  \{A \mid A \mbox{ is mapped to } \true \mbox{ under } I\}$ and $I^\bot = \{A \mid A \mbox{ is mapped to } \bot \mbox{ under } I \}$.
Atoms which do not occur in~$I^\true \cup I^\bot$ are mapped to~$\udf$.
The truth value of a given formula
under $I$ is determined according to the truth tables in
Table~\ref{tab:3vldluka}.
$I(F) = \top$ denotes that a formula $F$ is mapped to true under $I$.
A \defname{three-valued model} $\CalM$ of~$\CalP$ is a three-valued interpretation such that
$\CalM(A \leftarrow \Body) = \true$ for each $A \leftarrow \Body \in \Ground\CalP$.
Let $I = \langle I^\top, I^\bot \rangle$ and $J = \langle J^\top,
J^\bot \rangle$ be two interpretations.
$I \subseteq J$ iff $I^\top \subseteq J^\top$ and $I^\bot \subseteq J^\bot$. $I$ is the \textit{least model} of $\CalP$
iff for any other model~$J$ of $\CalP$ it holds that $I \subseteq J$.

\begin{table}[t]
\[\small
\begin{array}[c]{@{\hspace{0mm}}c|c}
F & \neg F\\ \midrule
\top & \bot \\
\bot & \top \\
\udf & \udf
\end{array}
\;\
\begin{array}{c|lll}
 \wedge & \top & \udf & \bot \\
\midrule
\top & \top & \udf & \bot \\
\udf & \udf & \udf & \bot \\
\bot & \bot & \bot & \bot
 \end{array}
\;\
\begin{array}{c|lll}
 \vee & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \top & \udf & \udf \\
\bot & \top & \udf & \bot
 \end{array}
\;\
\begin{array}{c|lll}
\leftarrow & \top & \udf & \bot \\
\midrule
\top & \top & \top & \top \\
\udf & \udf & \top & \top \\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|ccc}
  \leftrightarrow  & \top & \udf & \bot \\
  \midrule
\top & \top & \udf & \bot \\
\udf & \udf & \top & \udf\\
\bot & \bot & \udf & \top
 \end{array}
\;\
\begin{array}{c|c}
 L & \cxt(L)\\
\midrule
\top & \top \\
\bot & \bot  \\
\udf & \bot
\end{array}
\]
\caption{The truth tables for the connectives under the three-valued {\L}ukasiewicz logic and for $\cxt(L)$. $L$ is a literal, $\top$, $\bot$, and $\udf$ denote \textit{true}, \textit{false},
and \textit{unknown}, respectively.\label{tab:3vldluka}}
\end{table}

@@

\subsection{Reasoning with Respect to the Least Fixed Point of $\Svlp$}

For a given $\CalP$, consider the following transformation:
1.\ For each ground atom~$A$ which occurs as head of a clause in $\Ground \CalP$, replace
  all clauses of the form $A \leftarrow \Body_1, \ldots,$ $A \leftarrow
  \Body_m$ occurring in $\Ground \CalP$ by $ A \leftarrow \Body_1 \vee
  \ldots \vee \Body_m$. 2.\ Replace all occurrences of $\leftarrow$ by $\leftrightarrow$.
The obtained ground set of equivalences is called the \textit{weak completion}
of~$\CalP$ or~$\WC\CalP$.
Consider the following semantic operator, which is due to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
\rangle$ be an interpretation. $\Phi_\CalP (I) =
\langle J^\top,J^\bot\rangle$, where
\[
\begin{array}{lcl}
J^\top & = & \{ A \mid A \leftarrow \Body \in \Def(A,\CalP)
\mbox{ and
} \Body \mbox{ is \True under } \langle I^\top,I^\bot \rangle \} \\
J^\bot & = & \{ A \mid \Def(A,\CalP) \neq \emptyset \mbox{ and }\\
&& \rule{7.5mm}{0mm}\Body \mbox{ is \False under } \langle
I^\top,I^\bot \rangle \mbox{ for all }  A \leftarrow \Body \in \Def(A,\CalP)\}.
\end{array}\Shrink
\]
\cite{hk:2009a} showed that the
weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules, in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
always has a least model under  {\L}ukasiewicz logic,
which can be computed by the least fixed point of~$\Svlp$. The correspondence to other well-known logic programming approaches, in particular 
to the Well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.

The weak completion of contextual
programs might have more than one minimal model, and it is not guaranteed that 
the least fixed point of $\Svlp$ always exists. However,
if these programs do not contain cycles, then there always exists a supported model, which can be 
computed by iterating $\Svlp$: A least fixed point of $\Svlp$ always exists for contextual programs
that do not contain cycles~\cite{dietz:hoelldobler:pereira:2017}.
Let us denote the least fixed point of $\Svlp$ as $\CalM_\CalP$:
As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, we define
$\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
Note that for all programs that are considered in the following, a least fixed point of $\Svlp$ exists.

@@

\subsubsection{Conditional representation, license for inference and existential import}
Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
However, it seems that humans do not understand it as \textit{strict}, but defeasible~\cite{stenning:vanlambalgen:2008},
i.e.\ they implicitly assume a license for the inference (\licenses). For (\MA yz) this can be 
explicitly expressed by 
\textit{all $y$ that are not abnormal, are $z$}.
% \\
% Accordingly, $z(X)$ is conjoint with an abnormality predicate: 
% \[\forall X (y(X) \wedge \neg \Ab(X) \rightarrow z(X)).\] 
By default, we assume that \textit{nothing is abnormal}. % with respect to $X$}:
% \[
% \Ab(X) \leftarrow \bot.
% \]
Furthermore, humans have a pragmatic understanding of quantifiers in the language~\cite{grice1975}.
For instance, in natural language, we normally do not quantify over things that do not exist.
Consequently, \textit{All $z$} implies \textit{Some $z$ exist}.
This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import). 
It is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
 Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
require existential import for a conditional to be true.
$\CalP_{Ayz}$, the logic program representing (\MA yz) is as follows:
\begin{align} %{rll@{\hspace{\hfill}}r}
z(X) & \leftarrow  y(X) \wedge \neg \Ab_{yz}(X). 
\tag{\conditionals\ \&\ \licenses} \\
\Ab_{yz}(X) & \leftarrow \cxt(z'(X)). \tag{\licenses}\\
y(o) & \leftarrow \top. \tag{\import}
\end{align}
Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ can be introduced by the negation by transformation principle in the following.}
Nothing is abnormal, unless the negation of $z$ is true.
% , then $\Ab_{yz}(X)$ is false, for which then, given the first clause in $\CalP_{Ayx}$, $z(o)$ is true, if $y(o)$ is true. Note that, if $z'(X)$ would be written instead of $\cxt(z'(X))$, in the case where $\cxt(z'(o))$ is unknown, $z(o)$ could not be concluded, even if $y(o)$ would be true. This is due to the fact that negation in WCS is understood as strong negation.\footnote{This is different in other logic programming approaches, i.e.\ well-founded semantics.}
% \end{align}
% \]
We obtain \[\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle.\]

\subsubsection{Negation by transformation and no double negation}
